---
title: "Chr18_SnpEff"
author: "Dave Speca"
date: "7/22/2017"
output: html_document
---

####Preliminaries

Exploratory analysis of SnpEff output indicated that SnpEff ANN= data is appended to the INFO column of the .vcf file (column #8), making it quite unwieldy. But I should be able to use dplyr to search for specific strings (variant calls) that have been added by SnpEff.

####SnpEff analysis

N.B. I downloaded the latest version of SnpEff into my home directory. Then within the snpEff home directory I ran this command from the Terminal:

`java -Xmx4g -jar snpEff.jar mm10 Chr18/Chr18.all.vcf  > Chr18.all.ann.vcf`

This (quite rapidly) generated the Chr18.all.ann.vcf file.
N.B. Chr2.all.vcf was 18.9 Mb, Chr2.all.ann.vcf is 27.5 Mb, so this just adds additional columns of information. Then I started this new R_project (Chr18) after creating a git repository, and I moved the Chr18.all.ann.vcf file to the project's directory.

The SnpEff annotations are contained within the "INFO" variable. I will need to extract them.

```{r}
library(ggplot2)
library(dplyr)
```

###Read in the vcf file and add headers of Chr18.all.ann.vcf:

```{r, cache=TRUE}
# read in .vcf file
Chr18.data <- read.table("Chr18.all.ann.vcf", as.is=TRUE, na.strings=".")
vcf.header <- system("grep '#C' Chr18.all.ann.vcf",intern = TRUE)
vcf.header <- sub("#","",vcf.header) #get rid of the pound sign
vcf.header <- unlist(strsplit(vcf.header,split="\t"))
colnames(Chr18.data) <- vcf.header
head(Chr18.data)
```

###String split data

```{r}
# Before splitting add NAs to blank cells

Chr18.data$DBA[is.na(Chr18.data$DBA)] <- "NA:NA:NA:NA:NA:NA:NA:NA"

DBA.tmp <- matrix(
  unlist(strsplit(Chr18.data$DBA,split = ":")),
  nrow=nrow(Chr18.data),
  byrow=TRUE
  )
head(DBA.tmp)
colnames(DBA.tmp) <- paste("DBA",c("gt","tot.depth","NObsAllele", "ref.depth","ref.qual","alt.depth","alt.qual","gen.lik"),sep="_")

Chr18.data$Chr18[is.na(Chr18.data$Chr18)] <- "NA:NA:NA:NA:NA:NA:NA:NA"

Chr18.tmp <- matrix(
  unlist(strsplit(Chr18.data$Chr18,split = ":")),
  nrow=nrow(Chr18.data),
  byrow=TRUE
  )
head(Chr18.tmp)
colnames(Chr18.tmp) <- paste("Chr18",c("gt","tot.depth","NObsAllele", "ref.depth","ref.qual","alt.depth","alt.qual","gen.lik"),sep="_")

Chr18.data <- cbind(Chr18.data,DBA.tmp,Chr18.tmp,stringsAsFactors=FALSE)
summary(Chr18.data)
```


###Convert columns back to numeric:
```{r}
Chr18.data[,c("DBA_tot.depth","DBA_ref.depth","DBA_ref.qual","DBA_alt.depth","DBA_alt.qual",
            "Chr18_tot.depth","Chr18_ref.depth","Chr18_ref.qual","Chr18_alt.depth","Chr18_alt.qual")] <- 
  apply(Chr18.data[,c("DBA_tot.depth","DBA_ref.depth","DBA_ref.qual","DBA_alt.depth","DBA_alt.qual",
            "Chr18_tot.depth","Chr18_ref.depth","Chr18_ref.qual","Chr18_alt.depth","Chr18_alt.qual")],
        2,
        as.numeric
        )
head(Chr18.data, 6)
```

###Subset Chr18.data QUAL > 40

```{r}
# subset data keeping only SNPs with quality scores greater than or equal to 40
Chr18.Q40.data <- subset(Chr18.data, QUAL >= 40)
```

###Make it a little more user-friendly
```{r}
Chr18.Q40.tidy.data <-select(Chr18.Q40.data, -ID, -FILTER, -FORMAT, -DBA, -Chr18, -DBA_gen.lik, -Chr18_gen.lik, -DBA_NObsAllele, -Chr18_NObsAllele)
```




